---
title: "Workbook 1"
author: "Name"
date: "21 January, 2022"
output:
  word_document: default
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

First, let's create a code chunk and load some packages.

22

## Learning outcomes

1.  Explore R Studio environment
2.  Create/Edit/Knit Markdown workbooks
3.  Use packages
4.  Perform arithmetic operations and built-in functions and constants
5.  Learn about variables
6.  Learn about vectors

## 4. Arithmetic operations and built-in functions and constants

R understands the following basic operators:

-   \+  and - for addition and subtraction

-   \*   and / for multiplication and division

-   \^ for exponents

-   %% is the modulo operator

-   %/% for integer division

Run the following chunk and observe different answers (explain, why are
they different)

```{r}
1+7/3

1+7%%3

1+7%/%3

```

Below is an excerpt of some of the basic mathematical functions *R*
knows.

-   log() -- computes logarithms;

-   exp() -- computes the exponential function;

-   sqrt() -- takes the square root;

-   abs() -- returns the absolute value;

-   sin() -- returns the sine;

-   cos() -- returns the cosine;

-   tan() -- returns the tangent;

-   asin() -- returns the arc-sine;

-   factorial() -- returns the factorial;

-   sign() -- returns the sign (negative or positive);

-   round() -- rounds the input to the desired digit.

In addition to functions R has a small number of built-in constants:

-   LETTERS - the 26 upper-case letters of the Roman alphabet;

-   letters - the 26 lower-case letters of the Roman alphabet;

-   month.abb - the three-letter abbreviations for the English month
    names;

-   month.name - the English names for the months of the year;

-   pi - the ratio of the circumference of a circle to its diameter.

Run the following chunk

```{r}
LETTERS

sin(pi/2)

factorial(4)

sign(-125)

month.name

```

#### Exercise 1.4

Create a new chunk below this task and:

-   find the sum of squares of any 5 consequent integer numbers;

-   calculate the sum of the first 4 elements of MacLaurin series
    expansion for exponent function $e^x$ at the point x=2;

-   calculate $e^x$ at the point x=2 using R function *exp()* and
    compare to the previous result;

-   Let x be a number greater than 0. What are the results of the
    following operations?\
    Try to guess the answer at first, then evaluate these expressions by
    substituting x with any integer number\

    -   0/x\
    -   x/0\
    -   0/(x-x)
```{r}
q1<-1^2+2^2+3^2+4^2+5^2
paste( "5 consequent integer are 1 to 5,the sum of is:",q1)
#since any derivative of e^x is still e^x, the result is:
q2<-(exp(1))^0/factorial(0)*2^0+(exp(1))^0/factorial(1)*2^1+(exp(1))^0/factorial(2)*2^2+(exp(1))^0/factorial(3)*2^3
paste( "the sum is:",q2)
q3<-exp(2)
paste( "e^x at the point x=2 is:",q3,"         the difference between them is",q3-q2)
0/2
2/0
0/(2-2)

```

## 5. Variables

Briefly speaking, a variable defines a place in the computer memory
where you keep some value. Values could be of different type: numbers
(Integer, Floating point (Real), Long Decimal etc.), characters
(String), or logical (Boolean).

To assign a value to a variable in R you use an arrow, lie "\<-". The
arrow points from the data you want to store towards the name you want
to store it under. For example,

```{r}
a <- 8 + 3
```

If R has performed this command successfully you will not see any
output, as the value of 8 + 3 has been saved to the data structure
called *a*. However, you can see that x variable has appeared in the
Environment window. You can access and use this value at any time and
can print the value of a.

Run the following chunk, observe that there are only 3 values printed
(Why?)

```{r}
a
a^2
b <- a%/%3
log(x = 10, base = 10)
```

You can also display this value by using *print()* function:

```{r}
print(a)
print(b)
```

You can combine several commands on one line separated by semi-colon:

```{r}
print(a); print(b); a+b
```

Another useful symbol is *hash* (\#). It denotes beginning of the line
comments:

```{r}
exp(5)
round(exp(5),digits= 2)   # e^5 rounded to the nearest 0.01
round(exp(5),digits=-1)  # e^5 rounded to the nearest 10
```

If you do not know or do not remember the proper syntax of some
function, then you can either use Google or to use internal help by
typing question mark followed by the name of the function:

```{r}
?log()
```

For some functions, especially basic operators, ? may not work. In those
cases you can use the *help()* function:

```{r}
help("%/%")
```

#### Exercise 1.5

-   Create a new chunk.

-   Choose 4 numbers (integer or real), assign these values to variables
    named *a, b ,c, d* and calculate the following expressions:\
    $expression1=a+b \cdot c-d^d$\
    $expression2=a+(b \cdot c-d)^d$\
    $expression3=(a+b)\cdot c-d^d$\
    $expression4=a+b \cdot (c-d)^d$

```{r}
a <--1
b <-3.2
c <-4.1
d <-2
print("take a=-1,b=3.2,c=4.1,d=2.The result is:")
expression1=a+b*c-d^d
expression2=a+(b*c-d)^d
expression3=(a+b)*c-d^d
expression4=a+b*(c-d)^d

expression1
expression2
expression3
expression4

```


## 6. Vectors

There are several ways in R to create an array of values.
The easiest way to use the *c()* function. The “c???stands for concatenate, and you can string a bunch of elements
together, separated by commas.

```{r}
Vector1 <- c(1 ,2 ,3 ,4 ,5 ,6 ,7 ,8 ,9 ,10)

x <- c(65, 73, 54, 47, 90, 30, 44, 66, 77, 62)
```

You can perform arithmetic operations on all elemts.
For example, you can add a scalar (a number) to all elements or  square each entry:

```{r}
x + 2
x^2
```

#### Exercise 1.6.1

The weights of five people before and after a diet programme are given
as two arrays: Before: [78 72 78 79 105] After: [67 65 79 70 93] 

- Create a new chunk

- Create two vectors 'before' and 'after' and assign the corresponding values to them.  

- Calculate the difference between the values for each person.

```{r}
before <- c(78, 72, 78, 79, 105)
after <- c(67, 65, 79, 70, 93)
print("They lose weight respect:")
before-after

```


Some useful vectors can be created quickly. The colon operator (:) is used
to generate a sequences of integer values with step 1 or -1 depending on the first and the last values

```{r}
c1 <- 1:10
c1
c2<- -3:4
c2
c3<- 9:5
c3
```

Any arithmetic progression can be generated by *seq()* function

```{r}
seq(from=2, to=6, by=0.4)

seq(from=-1, to=1, length=6)
```

Sometimes it's necessary to have repeated values, for which we use *rep()*

```{r}
rep(5,3)

rep(2:5,each=3)
rep(-1:3, length.out=10)
```

We can also use R's vectorization to create more sequences:

```{r}
2^(0:10)

1:3 + rep(seq(from=0,by=10,to=30), each=3)
```

If we perform a binary operation (such as +) on two vectors of different
lengths, the shorter one is used over and over again until the operation
has been applied to every entry in the longer one - this is called
recycling.

If the longer length is not a multiple of the shorter length, a warning
is given.

```{r}
1:10 * c(-1,1)

1:7* 1:2 # this line will rise a warning
```

#### Exercise 1.6.2

Try to create the following vectors using rep() and seq() functions:

- 1, 1.5, 2, 2.5, . . . , 12;

- 1, 8, 27, 64, . . . , 1000; 

- 1, -1/2,1/3, -1/4, . . . , -1/100;

- 1, 0, 3, 0, 5, 0, 7, . . . , 0, 49.
```{r}
library(MASS)

print("Vector 1 is:")
seq(from=1, to=12, by=0.5)
print("Vector 2 is:")
(1:10)^(3)
print("Vector 3 is:")
fractions(1/rep(1:100, length.out=100)* c(1,-1))
print("Vector 4 is:")
seq(from=1, to=49, by=1)* c(1,0)
```

```{r, warning=FALSE, message=FALSE}
#library(tidyverse)
```


```{r, warning=FALSE, message=FALSE}
#library(readxl)
```
